Regular singular Volterra equations on complex domains

### 2023-08-16 reg-sing-volterra.tex
Show that kernel of V^alpha is meromorphic

To apply Theorem~\ref{thm:general_volterra}, we need to rewrite the equation $\hat{\mathcal{P}}_{\alpha}\varphi = 0$ in the form $\varphi = \volterra^\alpha \varphi$, where $\volterra^\alpha$ is a Volterra operator of the kind described in Section~\ref{setting}. \textcolor{orange}{[Explain how conditions shift from $\zeta = 0$ to $\zeta = \alpha$?]} When we write $\hat{\mathcal{P}}_\alpha$ explicitly as a Volterra operator, the equation $\hat{\mathcal{P}}_{\alpha} \varphi = 0$ becomes
\[ \big[ p\, \varphi \big](a) + \int_{\zeta=\alpha}^a q \, \varphi \; d\zeta + \int_{\zeta=\alpha}^a k_R(a,\cdot) \, \varphi \, d\zeta = 0, \]
where
\[ k_R(a,a')=\sum_{r=0}^{d-1}\sum_{j=0}^\infty \frac{R_{r,j}}{j!} \, \big(\zeta(a)-\zeta(a')\big)^{j+1} \, \big(-\zeta(a')\big)^r. \]
Isolating $p\,\varphi$ on the left and dividing both sides by $p$ yields the equivalent equation
\begin{equation}\label{eqn:divided}
\varphi(a) = -\int_{\zeta=\alpha}^a \frac{q}{p(a)} \, \varphi \; d\zeta - \int_{\zeta=\alpha}^a \frac{k_R(a,\cdot)}{p(a)} \, \varphi \, d\zeta.
\end{equation}
Let $\hardpart^\alpha$ be the Volterra operator with kernel
\[ \hardker^\alpha(a,a') = -\frac{q(a')}{p(a)}, \]
and let $\softpart^\alpha$ be the Volterra operator with kernel
\[ \softker^\alpha(a,a') = -\frac{k_R(a,a')}{p(a)}. \]
Now we can write Equation~\eqref{eqn:divided} as $\varphi = \volterra^\alpha \varphi$, where $\volterra^\alpha = \hardpart^\alpha + \softpart^\alpha$.
\par\color{RoyalBlue}
First of all, we rewrite $\hat{\mathcal{P}}_{\alpha}\varphi=0$ as $\varphi = \volterra_\alpha \varphi $: %we introduce the notation $p(a):=P(-\zeta(a))$ and $q(a)=Q(-\zeta(a))$, then $\hat{P}_j$ is

@@

\begin{align}
   \nonumber \varphi (a)& =-\int_{\zeta=\alpha}^a \frac{q}{p(\zeta(a))}  \, \varphi \, d\zeta' - \int_{\zeta=\alpha}^a \frac{k_R(a,\cdot)}{p(\zeta(a))}  \, \varphi \, d\zeta' \\
   \nonumber \varphi (a)& =\Big[ \hardpart^\alpha \, \varphi \Big](a)+ \Big[\softpart^\alpha \, \varphi\Big](a) \\
   \varphi(a) &=\Big[ \volterra^\alpha \, \varphi \Big](a)
\end{align}
\color{black}
\par\color{RoyalBlue}
Notice that $\hardpart^\alpha$ is holomorphic and separable, while $\softpart^\alpha$ is holomorphic and non-separable. In addition, $\hardpart^\alpha$ is of the type defined in Section \ref{setting:basic} and it satisfies Condition \eqref{cond:sing}: by Taylor expanding the polynomials $p$ and $q$ around $\zeta = \alpha$ \textcolor{orange}{[the shifted polynomial willhave the same degree as the original]} \textcolor{magenta}{[$\alpha$ is a number, not a point; write $c_0$ in place of $q(\alpha)$? notation $\zeta'$ not defined; switch to $\zeta(a')$?]}.
\begin{align*}
    -\frac{q}{p(a)}&=-\frac{\textcolor{magenta}{q(\alpha)}+c_1(\zeta\textcolor{magenta}{'}-\alpha)+c_2 (\zeta'-\alpha)^2+...}{-b_1(\zeta(a)-\alpha)+ b_2 (\zeta(a)-\alpha)^2+b_3(\zeta(a)-\alpha)^3+...}\\
    &=\frac{q(\alpha)}{b_1(\zeta(a)-\alpha)}\frac{1+c_1'(\zeta'-\alpha)+c_2' (\zeta'-\alpha)^2+...}{1+ b_2' (\zeta(a)-\alpha)+b_3'(\zeta(a)-\alpha)^2+...}\\
    &=\frac{q(\alpha)}{b_1(\zeta(a)-\alpha)}\frac{1+c_1'(\zeta'-\alpha)+c_2' (\zeta'-\alpha)^2+...}{1+b_2' (\zeta(a)-\alpha)+b_3'(\zeta(a)-\alpha)^2+...}
\end{align*}
hence setting $\tau_\alpha:=\textcolor{magenta}{-}\frac{q}{p'}\textcolor{orange}{\big|_{\zeta = \alpha}}\neq 0$, we find
\begin{equation*}
    -\frac{q(a')}{p(a)}-\frac{\tau_\alpha}{\zeta(a)-\alpha}
\end{equation*}
is bounded as $\zeta(a')$ and $\zeta(a)$ approach $\alpha$.
\color{black}
\par\color{SkyBlue}
Since $p$ has only simple roots, the partial fraction decomposition of $1/p$ is
\[ \sum_{\beta \in \roots} \left.\frac{1}{p'}\right|_{\zeta = \beta} \frac{1}{\zeta - \beta}, \]
where $\roots$ is the set containing the values of $\zeta$ at the roots of $p$.
\par\color{DarkTurquoise}
\textcolor{orange}{Nicer formulation; note different definition of $\roots$:}\par
Since $P(-\zeta)$ has only simple roots, the partial fraction decomposition of $1/P(-\zeta)$ is
\[ \sum_{-\beta \in \roots} \frac{1}{-P'(-\beta)\,(\zeta - \beta)}, \]
where $\roots$ is the zero set of $P$ \textcolor{orange}{[Ahlfors, {\em Complex Analysis}, \S1.4, Exercise~2]}. It follows that
\begin{align*}
-\frac{Q(-\zeta(a'))}{P(-\zeta(a))} & = \sum_{-\beta \in \roots} \frac{Q(-\zeta(a'))}{P'(-\beta)\,(\zeta(a) - \beta)} \\
& \in \sum_{-\beta \in \roots} \frac{Q(-\alpha) + O\big(|\zeta(a') - \alpha|\big)}{P'(-\beta)\,(\zeta - \beta)} \\
%%& \in \frac{Q(-\alpha) + O\big(|\zeta - \alpha|\big)}{-P'(-\alpha)\,(\zeta - \alpha)} + \sum_{\substack{-\beta \in \roots \\ \beta \neq \alpha}} \frac{Q(-\alpha) + O\big(|\zeta - \alpha|\big)}{-P'(-\beta)\,(\zeta - \beta)} \\
& \subset \frac{Q(-\alpha)}{P'(-\alpha)\,(\zeta - \alpha)} + O\big(1\big)
\end{align*}
at $(\zeta(a), \zeta(a')) = (\alpha, \alpha)$. In other words, for $\tau_\alpha = Q(-\alpha)/P'(-\alpha)$, the difference
\[ -\frac{Q(-\zeta(a'))}{P(-\zeta(a))} - \frac{\tau_\alpha}{\zeta - \alpha} \]
is bounded on a neighborhood of $\big(\zeta(a), \zeta(a')\big) = (\alpha, \alpha)$ in $\domain_\alpha^2$, showing that $\hardpart^\alpha$ satisfies Condition~\eqref{cond:sing}.

\textcolor{orange}{[Another phrasing:]}
It follows that
\begin{align*}
-\frac{Q(-\zeta(a'))}{P(-\zeta(a))} & = \sum_{-\beta \in \roots} \frac{Q(-\zeta(a'))}{P'(-\beta)\,(\zeta(a) - \beta)} \\
& = \sum_{\substack{-\beta \in \roots \\ \beta \neq \alpha}} \frac{Q(-\zeta(a'))}{P'(-\beta)\,(\zeta(a) - \beta)} + \frac{Q(-\zeta(a'))}{P'(-\alpha)\,(\zeta - \alpha)}
\end{align*}
As $a$ and $a'$ approach $\zeta = \alpha$, the
\[ \sum_{\substack{-\beta \in \roots \\ \beta \neq \alpha}} \frac{Q(-\zeta(a'))}{P'(-\beta)\,(\zeta(a) - \beta)} \]
part approaches the limit
\[ \sum_{\substack{-\beta \in \roots \\ \beta \neq \alpha}} \frac{Q(-\alpha)}{P'(-\beta)\,(\alpha - \beta)}. \]
Hence, that part is bounded on a neighborhood of $\big(\zeta(a), \zeta(a')\big) = (\alpha, \alpha)$ in $\domain_\alpha^2$. \textcolor{orange}{[\ldots]}

\begin{verify}
A kind of klunky way to do the Ahlfors exercise:
\begin{align*}
\frac{1}{p} & = \sum_{\beta \in \roots} \frac{\pi_\beta}{\zeta - \beta} \\
-\frac{p'}{p^2} & = -\sum_{\beta \in \roots} \frac{\pi_\beta}{(\zeta - \beta)^2} \\
p' & = \sum_{\beta \in \roots} \left(\frac{p}{\zeta - \beta}\right)^2 \pi_\beta \\
p'\big|_{\zeta = \alpha} & = \left.\left(\frac{p}{\zeta - \alpha}\right)^2\right|_{\zeta = \alpha} \pi_\alpha \\
& = \left[\prod_{\beta \in \roots \smallsetminus \alpha} (\alpha - \beta)\right]^2 \pi_\alpha \\
& = \left(p'\big|_{\zeta = \alpha}\right)^2 \pi_\alpha \\
\frac{1}{p'\big|_{\zeta = \alpha}} & = \pi_\alpha
\end{align*}
\end{verify}
\color{black}

>>>>

To apply Theorem~\ref{thm:general_volterra}, we need to rewrite the equation $\hat{\mathcal{P}}_{\alpha}\varphi = 0$ in the form $\varphi = \volterra^\alpha \varphi$, where $\volterra^\alpha$ is a Volterra operator of the kind described in Section~\ref{setting}. \textcolor{orange}{[Explain how conditions shift from $\zeta = 0$ to $\zeta = \alpha$?]} When we write $\hat{\mathcal{P}}_\alpha$ explicitly as a Volterra operator, the equation $\hat{\mathcal{P}}_{\alpha} \varphi = 0$ becomes
\[ \big[ p\, \varphi \big](a) + \int_{\zeta=\alpha}^a q \, \varphi \; d\zeta + \int_{\zeta=\alpha}^a k_R(a,\cdot) \, \varphi \, d\zeta = 0, \]
where
\[ k_R(a,a')=\sum_{r=0}^{d-1}\sum_{j=0}^\infty \frac{R_{r,j}}{j!} \, \big(\zeta(a)-\zeta(a')\big)^{j+1} \, \big(-\zeta(a')\big)^r. \]
Isolating $p\,\varphi$ on the left and dividing both sides by $p$ yields the equivalent equation
\begin{equation}\label{eqn:divided}
\varphi(a) = -\int_{\zeta=\alpha}^a \frac{q}{p(a)} \, \varphi \; d\zeta - \int_{\zeta=\alpha}^a \frac{k_R(a,\cdot)}{p(a)} \, \varphi \, d\zeta.
\end{equation}
Let $\hardpart^\alpha$ be the Volterra operator with kernel
\[ \hardker^\alpha(a,a') = -\frac{q(a')}{p(a)}, \]
and let $\softpart^\alpha$ be the Volterra operator with kernel
\[ \softker^\alpha(a,a') = -\frac{k_R(a,a')}{p(a)}. \]
Now we can write Equation~\eqref{eqn:divided} as $\varphi = \volterra^\alpha \varphi$, where $\volterra^\alpha = \hardpart^\alpha + \softpart^\alpha$.

The kernel of $\volterra^\alpha$ extends meromorphically over all of $\C^2$. One part of this observation demands some explanation: we need to confirm that the power series defining $k_R$ converges everywhere. To do this, recall that the functions $R_r(z^{-1})$ are all holomorphic on the disk $|z| > A$. Consequently, for any $\lambda > A$, we have $R_{r,j} \lesssim \lambda^j$ over all $r \in \{1, \ldots, d-1\}$ and all $j \in \{0, 1, 2, \ldots\}$. This is enough to show that $k_R$ is well-defined throughout $\C^2$. In fact, as we'll see later, it also leads to a useful bound on $k_R$.
\begin{verify}
\begin{align*}
& \sum_{j = 0}^\infty R_{r,j} z^{-j} \text{ converges when } |z| > A \\
\Longrightarrow & \sum_{j = 0}^\infty R_{r,j} z^{-j} \text{ converges when } |z^{-1}| < A^{-1} \\
\Longrightarrow & A^{-1} \le \operatorname{radius} R_{r,j} \\
\Longrightarrow & A \ge \big(\operatorname{radius} R_{r,j}\big)^{-1} \\
\Longrightarrow & A \ge \limsup_{j \to \infty} R_{r,j}^{1/j} \\
\Longrightarrow & A \ge \inf_{J \ge 0} \sup_{j \ge J} R_{r,j}^{1/j} \\
\Longrightarrow & \forall \lambda > A: \exists J: \sup_{j \ge J} R_{r,j}^{1/j} < \lambda \\
\Longrightarrow & \forall \lambda > A: \exists J: \forall j \ge J: R_{r,j}^{1/j} < \lambda \\
\Longrightarrow & \forall \lambda > A: \exists J: \forall j \ge J: R_{r,j} < \lambda^j
\end{align*}
\end{verify}

%%The least straightforward part of this observation is confirming that $k_R(a, a')$ is well-defined for all $a, a' \in \C$, which you can do by observing that

Since $P$ is a monic polynomial whose roots are all simple, $1/P(-\zeta)$ has the nice partial fraction decomposition
\[ \frac{1}{P(-\zeta)} = \sum_{-\beta \in \roots} \frac{1}{-P'(-\beta)\,(\zeta - \beta)}, \]
where $\roots$ is the zero set of $P$ \textcolor{orange}{[Ahlfors, {\em Complex Analysis}, \S1.4, Exercise~2]}. 
It follows that
\begin{align*}
-\frac{Q(-\zeta(a'))}{P(-\zeta(a))} & = \sum_{-\beta \in \roots} \frac{Q(-\zeta(a'))}{P'(-\beta)\,(\zeta(a) - \beta)} \\
& = \sum_{\substack{-\beta \in \roots \\ \beta \neq \alpha}} \frac{Q(-\zeta(a'))}{P'(-\beta)\,(\zeta(a) - \beta)} + \frac{Q(-\zeta(a'))}{P'(-\alpha)\,(\zeta(a) - \alpha)}
\end{align*}
As $a$ and $a'$ approach $\zeta = \alpha$, the
\[ \sum_{\substack{-\beta \in \roots \\ \beta \neq \alpha}} \frac{Q(-\zeta(a'))}{P'(-\beta)\,(\zeta(a) - \beta)} \]
part approaches the limit
\[ \sum_{\substack{-\beta \in \roots \\ \beta \neq \alpha}} \frac{Q(-\alpha)}{P'(-\beta)\,(\alpha - \beta)}. \]
Hence, that part is bounded on a neighborhood of $\big(\zeta(a), \zeta(a')\big) = (\alpha, \alpha)$ in $\domain_\alpha^2$. Conversely, as $a$ and $a'$ approach $\zeta=\alpha$, the term

\[\frac{Q(-\zeta(a'))}{P'(-\alpha)(\zeta(a)-\alpha)}\]

converges to $\frac{\tau_\alpha}{\zeta(a)-\alpha}$ with $\tau_\alpha:=Q(-\alpha)/P'(-\alpha)$. We conclude that 
\[-\frac{q(a')}{p(a)}-\frac{\tau_\alpha}{\zeta(a)-\alpha}\]

is bounded as $a,a'$ approach $\zeta=\alpha$, showing that $\hardpart^\alpha$ satisfies Condition \eqref{cond:sing}.

\begin{verify}
A kind of klunky way to do the Ahlfors exercise:
\begin{align*}
\frac{1}{p} & = \sum_{\beta \in \roots} \frac{\pi_\beta}{\zeta - \beta} \\
-\frac{p'}{p^2} & = -\sum_{\beta \in \roots} \frac{\pi_\beta}{(\zeta - \beta)^2} \\
p' & = \sum_{\beta \in \roots} \left(\frac{p}{\zeta - \beta}\right)^2 \pi_\beta \\
p'\big|_{\zeta = \alpha} & = \left.\left(\frac{p}{\zeta - \alpha}\right)^2\right|_{\zeta = \alpha} \pi_\alpha \\
& = \left[\prod_{\beta \in \roots \smallsetminus \alpha} (\alpha - \beta)\right]^2 \pi_\alpha \\
& = \left(p'\big|_{\zeta = \alpha}\right)^2 \pi_\alpha \\
\frac{1}{p'\big|_{\zeta = \alpha}} & = \pi_\alpha
\end{align*}

\color{magenta}
\begin{align*}
    \frac{1}{P} & = \sum_{\beta \in \roots} \frac{\pi_\beta}{\zeta - \beta} \\
-\frac{p'}{p^2} & = -\sum_{\beta \in \roots} \frac{\pi_\beta}{(\zeta - \beta)^2} \\
p' & = \sum_{\beta \in \roots} \left(\frac{p}{\zeta - \beta}\right)^2 \pi_\beta \\
&\\
& \\
P'\big|_{\zeta = \alpha} & = \left.\left(\frac{P}{t - \alpha}\right)^2\right|_{\zeta = \alpha} \pi_\alpha \\
& = \left[\prod_{\beta \in \roots \smallsetminus \alpha} (\alpha - \beta)\right]^2 \pi_\alpha \\
& = \left(p'\big|_{\zeta = \alpha}\right)^2 \pi_\alpha \\
\frac{1}{p'\big|_{\zeta = \alpha}} & = \pi_\alpha
\end{align*}
for some constant $c_\beta$.
\par\color{ForestGreen}
First, observe that
\begin{align*}
P(t) &= (t + \beta_1) \cdots (t + \beta_d) \\
P'(t) &= \sum_k \left[ \prod_{j \neq k} (t + \beta_j) \right] \\
P'(-\beta_1) &= \prod_{j\neq 1} (-\beta_1 + \beta_j) 
\end{align*}

\begin{align*}
\frac{1}{P(t)} & = \sum_{j} \frac{\pi_j}{t + \beta_j}\\
-\frac{P'(t)}{P(t)^2} &= -\sum_{j} \frac{\pi_\beta}{(t + \beta_j)^2}\\
P'(t) &= \sum_{j} \pi_j \left(\frac{P(t)}{t + \beta_j}\right)^2\\
P'(-\beta_1) &=  \pi_1 \prod_{j\neq 1} (-\beta_1+\beta_j)^2\\
\frac{1}{P'(-\beta_1)} &= \pi_1
\end{align*}
\end{verify}

\par\color{RoyalBlue}
First of all, we rewrite $\hat{\mathcal{P}}_{\alpha}\varphi=0$ as $\varphi = \volterra_\alpha \varphi $: %we introduce the notation $p(a):=P(-\zeta(a))$ and $q(a)=Q(-\zeta(a))$, then $\hat{P}_j$ is

@@

\begin{align}
   \nonumber \varphi (a)& =-\int_{\zeta=\alpha}^a \frac{q}{p(\zeta(a))}  \, \varphi \, d\zeta' - \int_{\zeta=\alpha}^a \frac{k_R(a,\cdot)}{p(\zeta(a))}  \, \varphi \, d\zeta' \\
   \nonumber \varphi (a)& =\Big[ \hardpart^\alpha \, \varphi \Big](a)+ \Big[\softpart^\alpha \, \varphi\Big](a) \\
   \varphi(a) &=\Big[ \volterra^\alpha \, \varphi \Big](a)
\end{align}
\color{black}
\par\color{RoyalBlue}
Notice that $\hardpart^\alpha$ is holomorphic and separable, while $\softpart^\alpha$ is holomorphic and non-separable. In addition, $\hardpart^\alpha$ is of the type defined in Section \ref{setting:basic} and it satisfies Condition \eqref{cond:sing}: by Taylor expanding the polynomials $p$ and $q$ around $\zeta = \alpha$ \textcolor{orange}{[the shifted polynomial willhave the same degree as the original]} \textcolor{magenta}{[$\alpha$ is a number, not a point; write $c_0$ in place of $q(\alpha)$? notation $\zeta'$ not defined; switch to $\zeta(a')$?]}.
\begin{align*}
    -\frac{q}{p(a)}&=-\frac{\textcolor{magenta}{q(\alpha)}+c_1(\zeta\textcolor{magenta}{'}-\alpha)+c_2 (\zeta'-\alpha)^2+...}{-b_1(\zeta(a)-\alpha)+ b_2 (\zeta(a)-\alpha)^2+b_3(\zeta(a)-\alpha)^3+...}\\
    &=\frac{q(\alpha)}{b_1(\zeta(a)-\alpha)}\frac{1+c_1'(\zeta'-\alpha)+c_2' (\zeta'-\alpha)^2+...}{1+ b_2' (\zeta(a)-\alpha)+b_3'(\zeta(a)-\alpha)^2+...}\\
    &=\frac{q(\alpha)}{b_1(\zeta(a)-\alpha)}\frac{1+c_1'(\zeta'-\alpha)+c_2' (\zeta'-\alpha)^2+...}{1+b_2' (\zeta(a)-\alpha)+b_3'(\zeta(a)-\alpha)^2+...}
\end{align*}
hence setting $\tau_\alpha:=\textcolor{magenta}{-}\frac{q}{p'}\textcolor{orange}{\big|_{\zeta = \alpha}}\neq 0$, we find
\begin{equation*}
    -\frac{q(a')}{p(a)}-\frac{\tau_\alpha}{\zeta(a)-\alpha}
\end{equation*}
is bounded as $\zeta(a')$ and $\zeta(a)$ approach $\alpha$.
\color{black}
\par\color{SkyBlue}
Since $p$ has only simple roots, the partial fraction decomposition of $1/p$ is
\[ \sum_{\beta \in \roots} \left.\frac{1}{p'}\right|_{\zeta = \beta} \frac{1}{\zeta - \beta}, \]
where $\roots$ is the set containing the values of $\zeta$ at the roots of $p$.
\par\color{DarkTurquoise}
\textcolor{orange}{Nicer formulation; note different definition of $\roots$:}\par

\color{black}
